\large{\bf Large effects and the infinitesimal model }

### 2023-12-04 writeup.tex
Fixing typos.

\paragraph{Reproductive Terms:}
We now turn our attention to the (point-wise) limit
\[
		\psi_{\infty}(\theta) = \lim_{n \to \infty} \psi_{n}(\theta).
\]
Since $\varphi(\theta)$ is a characteristic function, it has the Hermitian property: $\varphi(-\theta) = \overline{\varphi(\theta)}$, and thus $\psi_{n}(-\theta) = \overline{\psi_{n}(\theta)}$ for all $n$ and $\psi_{\infty}(-\theta) = \overline{\psi_{\infty}(\theta)}$ for all $\theta$ at which the limit exists.  We can thus, without loss, assume that $\theta > 0$.

Suppose $\psi_{\infty}(\theta)$ exists for some $\theta > 0$.  Then for all integers $k$, 
\[
	\psi_{\infty}(2^{k} \theta) = \lim_{n \to \infty} 2^{n}\psi\left(\frac{2^{k} \theta}{2^{n}}\right) =  \lim_{n \to \infty} 2^{k} 2^{n-k} \psi\left(\frac{\theta}{2^{n-k}}\right) 
	= 2^{k} \psi_{\infty}(\theta).
\]
In particular, since $\psi_{\infty}(0) = 0$, 
\[
	\frac{\psi_{\infty}(\theta)}{\theta} = \frac{\psi_{\infty}(2^{k} \theta)}{2^{k}\theta} = \frac{\psi_{\infty}(2^{k} \theta)-\psi_{\infty}(0)}{2^{k}\theta - 0}, 
\]
so if $\psi_{\infty}$ has a right-hand side derivative at zero, then
\[
	\frac{\psi_{\infty}(\theta)}{\theta} = \lim_{k \to -\infty} \frac{\psi_{\infty}(2^{k} \theta)}{2^{k}\theta} = \psi_{\infty}'(0+).
\]

Suppose that this derivative exists, say $\psi_{\infty}'(0+) = -\gamma + i m$ for some $m, \gamma \in \mathbb{R}$.  Then, by the Hermitian property,  $\psi_{\infty}'(0-) = \overline{\psi_{\infty}'(0+)}$ must exist as well, so
\[
	\psi_{\infty}(\theta) = \begin{cases} 
		i m\theta - \gamma \theta & \text{if $\theta \geq 0$, and}\\
		i m\theta + \gamma \theta & \text{if $\theta < 0$}.
	\end{cases}
\]
Moreover,  $\psi_{\infty}(\theta) = i m \theta - \gamma |\theta|$ is continuous at 0, and hence so is $\varphi_{\infty}(\theta) = e^{\psi_{\infty}(\theta)}$.  As we observed above, this implies that $e^{\psi_{\infty}(\theta)}$ is itself a characteristic function.  Thus, $e^{\psi_{\infty}(\theta)}$ must satisfy $|e^{\psi_{\infty}(\theta)}| = e^{\Re \psi_{\infty}(\theta)} \leq 1$, whence $\Re \psi_{\infty}(\theta) \leq 0$ and $\gamma \geq 0$.  In particular, if $\gamma > 0$, then we recognize $e^{\psi_{\infty}(\theta)} = e^{im \theta -\gamma|\theta|}$ as the characteristic function of a Cauchy distribution with median m and median absolute deviation $\gamma$ (\ie, the canonical $\alpha$-stable distribution with $\alpha=1$), \llabel{cauchy_note} whereas if $\gamma = 0$, then $e^{\psi_{\infty}(\theta)} = e^{im \theta}$ is the Fourier transform of a Dirac mass at $m$, \ie the characteristic function for a monomorphic population with all individuals having trait value equal to $m$.

>>>>

\paragraph{Reproductive Terms:}
We now turn our attention to the (point-wise) limit
\[
		\psi_{\infty}(\theta) = \lim_{n \to \infty} \psi_{n}(\theta).
\]

Suppose $\psi_{\infty}(\theta)$ exists for some $\theta \in \mathbb{R}$.  Then for all integers $k$, 
\[
	\psi_{\infty}(2^{k} \theta) = \lim_{n \to \infty} 2^{n}\psi\left(\frac{2^{k} \theta}{2^{n}}\right) =  \lim_{n \to \infty} 2^{k} 2^{n-k} \psi\left(\frac{\theta}{2^{n-k}}\right) 
	= 2^{k} \psi_{\infty}(\theta).
\]
In particular, since $\psi_{\infty}(0) = 0$, 
\[
	\frac{\psi_{\infty}(\theta)}{\theta} = \frac{\psi_{\infty}(2^{k} \theta)}{2^{k}\theta} = \frac{\psi_{\infty}(2^{k} \theta)-\psi_{\infty}(0)}{2^{k}\theta - 0}, 
\]
so if $\psi_{\infty}$ has a right-hand side derivative at zero, then
\[
	\frac{\psi_{\infty}(\theta)}{\theta} = \lim_{k \to -\infty} \frac{\psi_{\infty}(2^{k} \theta)}{2^{k}\theta} = \psi_{\infty}'(0+).
\]

Suppose that this derivative exists, say $\psi_{\infty}'(0+) = -\gamma + i m$ for some $m, \gamma \in \mathbb{R}$.  Since $\varphi(\theta)$ is a characteristic function, it has the Hermitian property: $\varphi(-\theta) = \overline{\varphi(\theta)}$, and thus $\psi_{n}(-\theta) = \overline{\psi_{n}(\theta)}$ for all $n$ and $\psi_{\infty}(-\theta) = \overline{\psi_{\infty}(\theta)}$ for all $\theta$ at which the limit exists.    In particular, if the right-hand side derivative, then $\psi_{\infty}'(0-) = -\overline{\psi_{\infty}'(0+)}$ must exist as well, so
\[
	\psi_{\infty}(\theta) = \begin{cases} 
		i m\theta - \gamma \theta & \text{if $\theta \geq 0$, and}\\
		i m\theta + \gamma \theta & \text{if $\theta < 0$}.
	\end{cases}
\]
(\nb we could have equally well assumed the left-hand side derivative existed and derived the right-hand side derivative).  Moreover,  $\psi_{\infty}(\theta) = i m \theta - \gamma |\theta|$ is continuous at 0, and hence so is $\varphi_{\infty}(\theta) = e^{\psi_{\infty}(\theta)}$.  As we observed above, this implies that $e^{\psi_{\infty}(\theta)}$ is itself a characteristic function.  Thus, $e^{\psi_{\infty}(\theta)}$ must satisfy $|e^{\psi_{\infty}(\theta)}| = e^{\Re \psi_{\infty}(\theta)} \leq 1$, whence $\Re \psi_{\infty}(\theta) \leq 0$ and $\gamma \geq 0$.  In particular, if $\gamma > 0$, then we recognize $e^{\psi_{\infty}(\theta)} = e^{im \theta -\gamma|\theta|}$ as the characteristic function of a Cauchy distribution with median m and median absolute deviation $\gamma$ (\ie, the canonical $\alpha$-stable distribution with $\alpha=1$), \llabel{cauchy_note} whereas if $\gamma = 0$, then $e^{\psi_{\infty}(\theta)} = e^{im \theta}$ is the Fourier transform of a Dirac mass at $m$, \ie the characteristic function for a monomorphic population with all individuals having trait value equal to $m$.
